---
title: "04 - 01 - Combining Data Sets"
output: html_document
date: "2023-10-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Concatenation

The simplest case of combining data is concatenation, in which multiple data 
sets are stacked atop one another or side-to-side. 

If the variables across the data sets to be stacked are the same, the rbind() function can be used to stack them. 

For example, to concatenate all three tournament data sets:

```{r}

# Manually created data set

dataset = data.frame(ID = c(1,2,3,4,5), Name = c("Alice", "Bob", "Carol", "David", "Eve"), Age = c(34, 48, 59, 50, 43), Distance = c(50, 30, 50, 50, 50))

# Imported CSV data set

dataset.csv <- read.csv("datafile.csv")

# Imported SAS7BDAT data set

install.packages("haven") # Need this again because projects on Posit Cloud are  
                          # entirely separated. Re-installing haven wouldn't be
                          # needed on a regular computer
library(haven)
dataset.sas <- read_sas("datafile.sas7bdat")

## Concatenating (stacking) the three data sets

dataset.rbind <- rbind(dataset, dataset.csv, dataset.sas)

```

Similarly, the cbind() function can be used to concatenate side-by-side. 

```{r}

dataset.cbind <- cbind(dataset, dataset.csv, dataset.sas)

```

The result of this is a data set with 12 columns and 5 rows. The first four c
olumns are the ID, Name, Age, and Distance variables from dataset because that 
was the first argument. The second four columns are those same variables from 
dataset.csv, and the third four columns are from dataset.sas. These had the 
same variable names, but R does not allow variables to have the same name 
within a data set. R automatically adjusted the variable names that appeared 
in columns 5 through 12: the variables from dataset.csv became ID.1, Name.1, 
Age.1, and Distance.1, and the variables from dataset.sas became ID.2, Name.2, 
Age.2, and Distance.2.

Straightforward concatenation is somewhat unusual in practice It is
much more common to combine data sets through merging.

## Merging one-to-one

A one-to-one merge works by first looking for observations in the data sets to
be merged that should be kept together because they share a source. This shared
source is sometimes referred to as a “primary key”, and it often takes the form 
of something like an ID number or a name.

The merge() function can be used to merge data one-to-one. It has three 
arguments. The first two are the names of the data sets to be merged. The third 
argument tells the merge() function on which variable the two data sets should 
be merged.

To demonstrate this, run the code below to create a new data object called
clubnames.

```{r}

clubnames <- data.frame(Name = c("Eve", "David", "Carol", "Bob", "Alice"), ClubName = c("School of Hard Nocks", "String Pullers", "Able Archers", "School of Hard Nocks", "Unaffiliated"))

```

The primary key shared between dataset and clubnames is Name, so that's the
variable we'll use to merge these. We don't need to sort them in advance.

```{r}

dataset.one.to.one <- merge(dataset, clubnames, by = "Name")

```

## Merging one-to-many

A one-to-many merge is very similar to the one-to-one merge in how it works; 
records from two data sets are combined based on a primary key. The difference
between the two is in the number of rows to be matched between the two 
constituent data sets. In a one-to-one merge, each observation in the first 
data set is matched to just one row from the second data set. In a one-to-many 
merge, each observation in the first data set is matched with more than one row 
from the second data set. 

To demonstrate this, run the code chunk below to import a new data set that 
contains Alice, Bob, Carol, David, and Eve's three most recent tournament
scores (i.e., 3 rows per participant):

```{r}

three.scores <- read.csv("PreviousThreeScores.csv")

```

The merge() function can be used to merge one-to-many using the primary key 
between data sets, which is again the Name variable.

```{r}

dataset.one.to.many <- merge(dataset.one.to.one, three.scores, by = "Name")

```

The new data set has 15 rows. The information from the five-row data set is 
repeated within participant, but the TotalScore values for each participant
differ based on their performance in the previous three tournaments. 

