---
title: "04 - 03 - Exercises Key"
output: html_document
date: "2023-10-31"
---

Run the following code chunk before starting the first question.

```{r}

first.committee = data.frame(SID = c(845675109, 687179692, 192023844, 231490119, 250596660, 414846148, 309019630), DistinctionRecommended = c("Yes", "Yes", "No", "No", "No", "Yes", "No"))

second.committee = data.frame(SID = c(949501458, 159099418, 470498231, 547650148, 516224505, 577887122, 509836207, 299120339), DistinctionRecommended = c("No", "No", "Yes", "No", "No", "Yes", "No", "Yes"))
```

1. Concatenate vertically (i.e., stack) these two data sets and save them as a 
new data set named recommendations.

```{r}

recommendations <- rbind(first.committee, second.committee)


```

Before starting the second question, run the following code chunk to import
a new data set. This data set contains equated scores for the students' 
Medicine, Pediatrics, and Surgery shelf exams. The students in this data set
are the same ones that are in the recommendations data set.

```{r}

exams <- read.csv("shelfexams.csv")

```

2. Merge the recommendations and exams data sets. What is the primary key 
shared by these data sets? Do you get the number of rows you'd expect based on 
your examination of the two data sets?

```{r}

# Primary key is SID

# Expect 45 rows in the merged data set (15 students in recommendations, 3 
# exam scores per student in exams)

committee.report <- merge(recommendations, exams, by = "SID")

```
