---
title: "02 - Functions, Packages, and Reading in Data"
author: "Wendy Christensen"
date: "2023-10-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Functions

Almost everything you'll want to do in R is done via a function. *Functions*
will always have a name followed by a set of parentheses. Functions generally
require that you enter at least one *argument* into the parentheses. 

You already used a function in the last unit: the abs() function.

All functions have documentation, which can be viewed by typing ? before the
name of the function. To examine the documentation of the abs() function, run
the code chunk below. 

```{r}

?abs()

```

From the documentation that appeared in the Help panel on the bottom-right,
the abs() function computes the absolute value of a single parameter. As you saw
in the last unit, abs() will accept a number directly as an argument and will
also accept a saved value that contains a number. 

If a function has more than one argument, separate entries with a comma.

## Creating Data

The most simple form of data is a single row/column of values, called a 
*vector*. You can manually create a vector in R by using the c() function. 
The 'c' stands for concatenate. Each element you input needs to be separated
by a comma. 

```{r}

B = c(1,2,3,4)

B

```

After running the chunk, there is a a new value in the Environment panel named 
a with the four elements that were entered into the c() function.

Most data sets contain both rows and columns, and the columns usually have 
names. One way to manually create a data set with rows and columns in R is to 
use the data.frame() function. The data.frame() function will take any number 
of vectors and add them to the data set as columns. The vectors need to be 
separated by commas, and it's important that the parentheses for both the c() 
and data.frame() functions are correctly closed. 

```{r}

dataset = data.frame(ID = c(1,2,3,4,5), Name = c("Alice", "Bob", "Carol", "David", "Eve"), Age = c(34, 48, 59, 50, 43), Distance = c(50, 30, 50, 50, 50))

```

To view this new data object, click on "dataset" in the Environment panel. A
new window will appear showing the data set.

To view the variable types, click on the blue circle with a white triangle 
next to "dataset". The variable names, their types, and the first few values
will be displayed under "dataset". To hide these, click on the blue circle
with a white triangle again. 

## Reading in data from CSV files

A common file type for external data sets is the comma-delimited file (CSV). 
Part of the reason for the popularity of CSV files is that most statistical 
software, including R, can easily import them. 

To import a CSV file as a data set, use the read.csv() function. The only
argument needed is the file name (including the file extension) in quotes. By
default, the read.csv() function use the first row as column names. 

```{r}

dataset.csv <- read.csv("datafile.csv")

```

## Reading in data from SAS7BDAT files using the haven package

Sometimes one needs to read in data files created by other statistical 
programs. One example of such a file is the SAS7BDAT file, which is native to 
SAS. R does not have a function built-in to read SAS7BDAT files, but R allows 
you to install *packages* that contain functions that other users have created 
and made available to share with other users. Packages must be installed before
any functions in them can be used. 

Packages can be installed in a few different ways. In this HOW, we'll use
the install.packages() function. You must know the exact name of the package. 
Remember that everything is case sensitive!

Run the code chunk below to install the haven package. 

```{r}

install.packages("haven")

```

Lots of red text appeared in the Console, but don't worry - nothing has gone 
wrong. The red text in this case is more like a log showing the steps R took to 
install the package. Many packages use functions from other packages as part of
their functionality; these other packages are called *dependencies*. When you 
tell R to install a package, it will also install any dependencies of the 
package. The haven package uses functions from many other packages, which is 
why there is so much red text. 

You only need to install a package once, but you must load the package into 
memory every time you start a new R session. 

Run the code chunk below to load the haven package. Once run, you can use
functions from a package any time during an R session. 

```{r}

library(haven)

```

The function from the haven package we need to import SAS7BDAT files is the 
read_sas() function. Like the read.csv() function, the only argument needed
is the filename (with the extension) in quotes.

```{r}

dataset.sas <- read_sas("datafile.sas7bdat")

```


