---
title: "02 - 03 - Exercises Key"
author: "Wendy Christensen"
date: "2023-10-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What's missing? 

Run the code in the code chunk below, which will return an error message. There
is one mistake in the code chunk; find it and correct it. If your change is
correct, the code chunk will run without error and a new data set named fixme
will appear in the Environment panel. 

```{r}

# The mistake is that the last right parenthesis is missing. 

fixme = data.frame(ID = c(1,2,3,4), Height = c(62, 70, 65, 71), EyeColor = c("Blue", "Brown", "Brown", "Brown"), Age = c(36, 44, 52, 41))

```


## Reading in a CSV file

There is a CSV file in the project folder named import.csv. Create a code 
chunk below for your code, then use the read.csv() function to to read in the 
data set and name the data set exercise.csv. 

```{r}

exercise.csv <- read.csv("import.csv")

```


## Reading in a SAS7BDAT file

There is an SPSS file in the project folder named import.sav. Create a code 
chunk below for your code, load the haven package, then use the read_spss() 
function to read in the data set and name the data set exercise.spss . 

```{r}

exercise.spss <- read_sav("import.sav")

```