---
title: "03 - 03 - Exercise Part 1 Answers"
author: "Wendy Christensen"
date: "2023-10-31"
output: pdf_document
---

Run the following code chunk before completing the exercise.

```{r}

registration = data.frame(Registrant = c(100,101,102,103,104,105,106,107,108,109)
, RegType = c("Student", "Student", "Presenter", "Presenter", "Regular", "Regular", "Regular", "Regular",  "Regular", "Presenter") , HotelReserved = c("Yes", "No", "No","No","Yes","Yes","Yes","No","Yes","Yes"), RegPayment = c("Card", "Card", "Check","Card","Check","Check","Card","Card","Card","Card"))

```

Do the following steps in order:

1. You learn that Registrant 103 actually does have their hotel room reserved.
Create a new data set named registration2 and change Registrant 103's value
for HotelReserved to "Yes".

```{r}

# Other solutions exist, but here's one that will work

registration2 <- registration

registration2[4,3] <- "Yes"

```

2. You learn that Registrant 108 will not be attending. Remove this registrant
from the registration2 data set. 

```{r}

registration2 <- registration2[-9,]

```

3. For some reason, it's bothersome that Registrant has a numeric variable 
type. Change this variable to a character type variable in the registration2
data set.

```{r}

registration2$Registrant <- as.character(registration2$Registrant)

```
