---
title: "03 - 01 - Manipulating Data Sets"
author: "Wendy Christensen"
date: "2023-10-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1

## Using row-and-column notation to add/remove rows and columns

The code chunk below creates the data set we manually created in the last unit.

```{r}

dataset = data.frame(ID = c(1,2,3,4,5), Name = c("Alice", "Bob", "Carol", "David", "Eve"), Age = c(34, 48, 59, 50, 43), Distance = c(50, 30, 50, 50, 50))

```

Because data files usually consist of rows and columns, it’s natural to think 
of this data structure as a matrix. Matrices have a convenient row-and-column 
notation style (rows x columns) that makes it easy to identify any element, 
any row, or any column in a matrix. R allows users to use row-and-column 
notation to identify and perform operations on the elements, rows, columns of a 
data set. Brackets indicate that the notation is being used, and a comma is
used within the brackets to separate rows (left side of comma) and columns 
(right side of comma).

Below is an example of code that pulls individual rows, columns, and elements
from dataset. The Name columns is the second column, the row containing 
David's information is in row 4, and David's name is in the fourth row of
column 2 (4 x 2).

```{r}

Name.col = dataset[,2] # Selects Name column (column 2)
David.row = dataset[4,] # Selects David’s row (row 4) 
David.name = dataset[4,2] # Selects David’s name (row 4, column 2)

```

To remove a column from a data object, this notation can be combined with a 
minus sign to indicate that the column should be dropped. For example, the code
below first removes the ID column from dataset then removes David's row. To
commit the changes to the original data object, the assignment operator can be
used to overwrite the existing data object to reflect the desired change.

```{r}

dataset <- dataset[,-1] # Removes ID column

dataset <- dataset[-4,] # Removes David's row column

```

If you re-examine the data set, you'll see that there are now three columns and
four rows. If you want to preserve the original data object while still making
changes, you can save the desired change as a new data object.

To add a column or a row, a vector with the appropriate number of elements can 
be added as a new row or a new column using the rbind() and the cbind() 
functions, respectively. The arguments for these functions are the data objects 
or vectors that one wants to bind together, each of which is separated by a 
comma. If the goal is to add something to an existing data set, then the result 
should be assigned to the same name as the original data set. 

For example, if I wanted to add a new entry to the data set with the Name, 
Age, and Distance columns, I could create a vector with the new information and 
use the rbind() function to add the row like so:

```{r}
 
new.entry <- c("Zander", 51, 30)  # Creates Zander's vector

dataset <- rbind(dataset, new.entry) # Adds Zander to dataset

```

If we re-examine dataset, we see that Zander's information has been added as a 
new row. 

## Setting variable types

But something else has changed about the data set. If we examine the variables 
in the Environment panel, we see that all the variables are now character v
ariables. This will cause problems later if we want to perform numeric 
operations on the columns that contain numbers (e.g., taking a mean of Age).

To change a variable to a numeric or character variable type, use the 
as.numeric() and as.character() functions, respectively. These functions can be 
combined with the row-and-column notation discussed in the previous section to 
target specific variables in an existing data set. 

Name should be a character variable, so that one does not need to be changed.
Age and Distance should be numeric, so the as.numeric() function can be used
to change these to numeric variables. 

```{r}

dataset[,"Age"] <- as.numeric(dataset[,"Age"]) 

dataset[,"Distance"] <- as.numeric(dataset[,"Distance"])

```

Notice that the typing has been changed in the Environment panel. 


# Part 2

## Creating new variables based on existing variables

Another common data manipulation task is to create new variables based on the 
values of existing variables. One way to do this is to use the ifelse() 
function, which takes three arguments. The first argument is the condition on 
which the new variable should be created, the second is what value the new 
variable will take if the condition is true, and the third is what value the 
new variable will take if the condition is false.

Many different types of conditional statements can be made by using one or more 
comparison operators. This is list of the comparison operators in R: 
== (equal to) 
!= (not equal to) 
> (greater than) 
>= (greater than or equal to) 
< (less than) 
<= (less than or equal to)

If one needs to create a new variable based on multiple conditions, logical
operators can be used. Some common ones are AND, OR, and NOT. In R, AND is 
represented by & (the ampersand key), OR is represented by | (the pipe key), 
and NOT is represented by ! (the exclamation point key).

In it's most straightforward version, the ifelse() function creates a new
binary variable based on one condition operator applied to one exiting 
variable. For example, let's say I want people who are less than 50 years old 
to be considered “Senior” and people 50 or more years old to be “Master”. The
first argument in the function is the condition statement (Age < 50), the
second argument is the value the new variable will take if the condition is
true (Senior), and the third argument is the value the new will take if the 
condition is false (Master). 

```{r}

dataset$AgeLevel <- ifelse(dataset$Age < 50, "Senior", "Master")

```

There is now a new variable named AgeLevel in dataset, such that Alice, Bob, 
and Eve are Senior and Carol and Zander are Master. 

To have three or more categories, ifelse() functions can be nested to allow 
for as many condition evaluations as needed. The total number of ifelse() 
functions needed to create a new categorical variable is one fewer than the 
desired number of categories. Recall that the ifelse() function has three 
arguments: the conditional statement, the resulting value if the condition is 
true, and the resulting value if the condition is false. Thus, the appropriate 
place to nest suceeding ifelse() functions is in that third argument. R will
evaluate each one in order of appearance and will stop once a condition is met.

Let’s say that I want to create a four-category variable, Division, based on 
the values of Age and Distance. The values Division should take are “Senior 
30m”, “Senior 50m”, “Master 30m”, and “Master 50m”. A combination of comparison 
operators, logical operators, and nested ifelse() functions will let me do 
this. Because I want to make four categories, I need three ifelse() statements.

```{r}

dataset$Division = ifelse(dataset$Age < 50 & dataset$Distance == 30, "Senior 30m", 
                          
                   ifelse(dataset$Age < 50 & dataset$Distance == 50, "Senior 50m",
                          
                   ifelse(dataset$Age >= 50 & dataset$Distance == 30, "Master 30m", "Master 50m")))

# The extra white space is to make this easier to read

```

* The AND (&) logical operator is used in the first argument in all three 
ifelse() functions to have Age and Distance evaluated simultaneously in the
condition statement; the condition is fulfilled only if both parts of it are
true. 

* The first ifelse() function will set the new variable to “Senior 30m” when 
both conditions in its first argument are met (Age less than 50 and Distance is 
30). The third argument of the first ifelse() function contains the second 
ifelse() function, which is used if the condition is not met. 

* The second ifelse() function will set the new variable to “Senior 50m” when 
both conditions in its first argument are met (Age less than 50 and Distance is 
50). The third argument of the second ifelse() function contains the third 
ifelse() function, which is used if the condition is not met. 

* The third and final ifelse() function will set the new variable to “Master 
30m” when both conditions in its first argument are met (Age greater than or 
equal to 50 and Distance is 30). If this condition is not met, it will set
the new variable to "Master 50m".

## Subsetting data

Subsetting data refers to selecting a subset of observations or a subset of 
variables in a data set. In the case of subsetting variables, the 
row-and-column notation discussed in a previous section can be expanded to 
keeping or removing multiple variables. When variables are adjacent, colon 
notation can be used to save some typing.

```{r}

first.three.vars <- dataset[,c(1:3)] # Contains Name (col 1), Age (col 2), and                                           
                                     # Distance (col 3)

first.last.vars <- dataset[,c(1,5)] # Contains Name (col 1) and Division (col 5)

middle.three.vars <- dataset[,-c(1,5)] # Contains Age (col 2), Distance (col 3),
                                       # and AgeLevel (col 4)

```

A similar process can be used to keep or remove rows

```{r}

first.three.obs <- dataset[c(1:3),] # Contains Alice (row 1), Bob (row 2), and
                                    # Carol's (row 3) rows

first.last.obs <- dataset[c(1,5),] # Contains Alice (row 1) and Zander's (row 3)
                                   # rows

middle.three.obs <- dataset[-c(1,5),] # Contains Bob (row 2), Carol (row 3), 
                                      # and Eve's (row 4) rows

```

It is also common to subset rows based on their contents. One way to do this is
the subset() function, which has two arguments, The first argument is the name 
of the data set to be subsetted, and the second argument is the condition 
statement describing which rows should be kept. 

```{r} 

Seniors50m <- subset(dataset, AgeLevel == 'Senior' & Distance == 50) 

```

The result contains just Alice and Eve, the only two who meet both conditions.


