---
title: "05 - 01 - Summarizing Data"
output: pdf_document
date: "2023-10-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summarizing continuous variables

Creating summaries of continuous variables usually involves summary statistics 
about the central tendency of the data (e.g., the mean and median) and the 
variability of the data (e.g., the minimum value, the maximum value, the first
and third quartiles, the interquartile range, standard deviance, and variance)

From the data set that was created through a one-to-one merge in the last
module, dataset.one.to.one, we can compute these statistics for the continuous 
Age variable.

```{r}

dataset.one.to.one <- read.csv("data.one.merged.csv")

mean(dataset.one.to.one$Age) # Returns 46.8 

median(dataset.one.to.one$Age) # Returns 48 

min(dataset.one.to.one$Age) # Returns 34 (Alice’s age) 

max(dataset.one.to.one$Age) # Returns 59 (Carol’s age) 

quantile(dataset.one.to.one$Age, 0.25) # Returns 43 

quantile(dataset.one.to.one$Age, 0.75) # Returns 50 

IQR(dataset.one.to.one$Age) # Returns 7 

sd(dataset.one.to.one$Age) # Returns 9.20326 

var(dataset.one.to.one$Age) # Returns 84.7

```
Through packages, R has incredible capacity for producing graphics; I highly 
recommend the R Graphics Cookbook (https://r-graphics.org) if you want to get a
sense for the variety of visualizations R can do. When getting to know one’s 
data, though, quick and simple functions are useful because of their simplicity. 

The plot() function can be used to produce a simple scatter plot of a single 
continuous variable.

```{r}

plot(dataset.one.to.one$Age)

```

The plot would be more interesting if there were more than five observations, 
but it gives a good visual sense for the statistics that were computed 
previously. To plot two continuous variables, add the second continuous 
variable (separated by a comma) as a second argument to the plot() function.
The first argument in the plot() function is set as the Y-axis, and the second
is set as the X-axis. When only one argument is included as above, that single
variable is plotted on the Y axis and an Index, corresponding to the order 
of the observations in the data set, is added to the X-axis.

Another option for visualizing a continue variable is a histogram. The hist()
function requires just one argument, the variable to be plotted as a histogram.

```{r}

hist(dataset.one.to.one$Age)

```
Yet another option is a boxplot, for which the boxplot() function can be used. 

```{r}

boxplot(dataset.one.to.one$Age)

```

We know from the summary statistics that the median is 48, which is where the 
bolded line is. The first and third quartiles, 43 and 50 respectively, are the 
edges of the box. The whiskers of the plot represent the values that are less 
than the first quantile and greater than the third quantile.



## Summarizing categorical variables

A common method of summarizing categorical variables is the frequency table, 
which displays counts and/or proportions of the number of observations in 
different categories. One way to look at frequencies in R is the table() 
function. The table() function requires just one argument, which will produce a 
row of counts for a single categorical variable. One such variable in our
merged data set is ClubName.

```{r}

table(dataset.one.to.one$ClubName)

```

From this table, we can see that two participants identified their club as 
“School of Hard Nocks” and one participant each identified their club as “Able 
Archers” and “String Pullers”, and one participant reported no club affiliation.

Contingency tables, also called cross-tabulation tables, are frequency tables 
in which two categorical variables are crossed. The levels of one variable make
up the rows, and the levels of the other variable make up the columns, and 
counts appear wherever both conditions in a cell apply to an observation. When 
using the table() function to create a contingency table, the first argument 
designates the rows and the second designates the columns.

```{r}

Club.Distance.table <- table(Club = dataset.one.to.one$ClubName, "Distance in M" = dataset.one.to.one$Distance)

Club.Distance.table 

```
In addition to adding Distance as a second variable, this code also 
demonstrates a convenient option for labeling the axes of the contingency 
table. The row axis label, Club, did not require quotes because it was a single 
word. The column axis label, Distance in M, needed quotes because the label was 
more than one word.

It is sometimes useful to see proportions of a total rather than raw counts. 
In R, the proportions() function computes the proportion of the total for each 
cell in the contingency table.

```{r}

proportions(Club.Distance.table)

```

