---
title: "05 - 03 - Exercise Answers"
output: pdf_document
date: "2023-10-31"
---

Run the following code chunk before starting the exercises.

```{r}

orders = data.frame(OrderTotal = c(1054, 304, 822, 1276, 998, 325, 620, 844, 902, 1099, 499, 914, 850, 723, 999), OrderType = c("Set", "Parts", "Parts", "Set", "Set", "Parts", "Parts", "Parts", "Set", "Set", "Parts", "Set", "Parts", "Parts", "Set"), OrderSource = c("Website", "Website", "Store", "Store", "Store", "Website", "Website", "Store", "Website", "Store", "Store", "Website", "Website", "Website", "Store"))

```

1. Compute the mean and standard deviation of the OrderTotal variable.

```{r}

mean(orders$OrderTotal)
sd(orders$OrderTotal)

```

2. Plot the OrderTotal variable as a histogram.

```{r}

hist(orders$OrderTotal)

```

3. Create a frequency table for the OrderType variable. Which type of order
was more common?

```{r}

table(orders$OrderType) # 8 orders were Parts orders and 7 were Sets; parts
                        # orders are more common.

```

3. Create a contingency table for the OrderType and OrderSource variables. 
Which type of order was more common? Label the axes as "Order Type" and "Order 
Source". Which combination was most common?

```{r}

table("Order Type" = orders$OrderType, "Order Source" = orders$OrderSource)

# Most common combination was parts orders from the website

```