---
title: "01 - 01 - Basic Syntax and the RStudio Environment"
author: "Wendy Christensen"
date: "2023-10-31"
output:
  pdf_document: default
---

## Using the Console

The Console is where code you write is executed and sometimes printed line by 
line. After you've completed a line, you press Enter/Return to execute the line. 
Try typing each of the lines below into the Console:

1 + 2 

abs(-100) 

You can assign values to named objects. You can do this by using an assignment
operator. In most cases, you can use either = or <- to assign values to 
objects, but using the latter is considered better practice because there are 
some edge cases where the two produce different results. Try typing each of the 
lines below into the Console and look at what appears in the Environment panel 
(top right window):

a = 1 
       
b <- 2 
       
c <- 4^4

c

If an operation is directly typed into the Console, R will perform it and
print the result in the Console. If the operation is saved as a value, it will
not print in the Console but will appear as a new value in the Environment 
panel. 

Try typing the next two lines below into the Console:

d <- a + b 
       
D <- c/2 

Notice how d and D appear as separate values in the Environment panel. This is
because R is case sensitive - d and D are not the same! In addition to having 
correct syntax, you must also check that the case of what you've typed is 
correct. 

To clear the saved values, press the broom icon above the Environment 
panel. If you clear the saved values, you will need to redefine them by 
re-running their code if you want to use them again. You will see an error 
message if you try to use a saved variable that has been cleared.

***

## R Markdown

R Markdown files are a special type of file that allows for word processing and
code execution in the same document. R Markdown files, which have the .rmd file 
extension, use Markdown to add formatting to text. There are many flourishes 
one can add, two of which are shown in this document: headers (##) and line 
breaks (***). A cheat sheet produced by Posit (makers of RStudio) is in the 
project documents in this unit. 

## R Markdown

Code is written and executed within "chunks". All code chunks start with 
```{r} and end with ```. You will know you've successfully set up a code 
chunk when the background turns grey and chunk controls appear at the top-right 
of the chunk.

```{r}

# A pound sign creates a code comment, which is colored green - very useful! 

# In code chunks, you can execute single lines by setting your cursor on a line
# and pressing Control + Enter

x <- 10

# Multiple lines can be executed by highlighting multiple lines and pressing
# Control + Enter

y <- 5
z <- x*y

# All code in a chunk can be executed by pressing the play button at the
# top-right of the chunk

```

R Markdown is also useful for sharing writing and code in other file formats. 
To export this document as a PDF file, select Knit -> Knit to PDF. The compiled 
PDF file will appear in the Files panel. 

